RationalGRL: A Framework for Argumentation and Goal Modeling

### 2017-01-26 background.tex
example

\subsection{Goal-oriented Requirements Language (GRL)}
\label{sect:background:grl}
GRL is a visual modeling language for specifying intentions, business goals, and non-functional requirements (NFRs) of multiple stakeholders \cite{Amyot:2010:EGM:1841349.1841356}. It can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions. A GRL model is a connected graph of intentional elements that optionally are part of actors. Figure~\ref{fig:trafficsim} illustrates a GRL diagram. An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system (\textbf{User}, Figure~\ref{fig:trafficsim}), or the system itself (\textbf{Simulation}, Figure~\ref{fig:trafficsim}). Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied. Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way. Softgoals (e.g. \textbf{Realistic simulation)} are often more related to NFRs, whereas goals (such as \textbf{Generate cars}) are more related to functional requirements. Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals. In Figure~\ref{fig:trafficsim}, some of the tasks are \textbf{Create new cars} and \textbf{Keep same cars}. In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g. \textbf{******})

@@

Different links connect the elements in a GRL model. AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements. In Figure~\ref{fig:trafficsim}, the goal \textbf{Generate cars} is XOR-decomposed to the tasks \textbf{Create new cars} and \textbf{Keep same cars}. Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element. A contribution link has a qualitative contribution type or a quantitative contribution. Task \textbf{Create new cars} has a \textbf{help} qualitative contribution to the softgoal \textbf{Realistic simulation}. Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors. For example, actor \textbf{Simulation} depends on the actor \textbf{User} to perform the task \textbf{Adjust car spawning rate} to fulfil its task \textbf{Create new cars}.

@@

In this paper, we use the classic example of an online meeting scheduler that has been used in the literature~\cite{}, ~\cite{}, ~\cite{} several times. The meeting scheduler has been modeled with Tropos~\cite{} and $i*$. We remodel this example with GRL in jUCMNav for the purpose of our framework.

The meeting scheduler system has three actors, \textbf{Meeting Initiator}, \textbf{Meeting Scheduler} and \textbf{Meeting Participant}. The \textbf{Meeting Scheduler} has a goal that for every meeting request, it has to try to find the best time and location in a way that most of the intended participants (i.e. \textbf{Meeting Participant}) can attend the meeting. The \textbf{Meeting Initiator} organizes the meetings by asking the participants to give their availability for a range of dates and time as well as the dates that they are not available. Based on these information, the \textbf{Meeting Scheduler} needs to pick a date/time and a location from the available dates/times given by the participants in a way that the most number of participants can attend. Participants should then accept or reject the event.

@@

\todo{Sepideh}{Marc}{I think you can add some more or remove/modify some of these based on your needs in the future sections}
\begin{itemize}
	\item Why are only the two softgoals \textbf{Quick} and \textbf{Low effort} selected for the task \textbf{Organize Meeting}? Why is, for example, a goal \textbf{Selecting the most convenient time} not included in the analysis for the actor \textbf{Meeting Initiator}?
	\item What does \textbf{Quality of Proposed Date} mean?
	\item How can one decide if the quality is low or high? Who is in charge? \todo{Floris}{Sepideh}{do you mean the \textbf{Quality of Proposed Date}? Or quality of the GRL diagram?}
	\item Does the meeting initiator use an email to inform the meeting participants to enter their availabilities? 
	\item Does the meeting scheduler uses an online calendar to obtain the available dates or an online website?
	\item How much time do the participants have to enter their availabilities? Is that important? Do they get a reminder from either the initiator or the scheduler? Is the reminder sent via an email? 
	\item What does \textbf{Richer Medium} mean? 
\end{itemize}

\todo[resolved]{Sepideh}{all}{We need to further improve this part}
\todo{Marc}{all}{Improved this part, what do you think?}
These are the type of the questions that we cannot answer just by looking at the GRL models. The model in figure~\ref{fig:trafficsim} does not contain information about discussions that let up to the resulting elements of the model, such as by various clarification steps for the naming, or alternatives that have been considered for the relationships. In order to address this, we use Argument Schemes for Goal Modeling (GMAS). GMAS can help us decide what intentional elements should remain in GRL and what new ones to be added or deleted, provide rationale behind the design decisions and the relationships between the links.

@@

So, for example, we can say that 
\begin{itemize}
\item[] \textbf{Find agreeable meeting date} is a goal,
\item[] \textbf{Find agreeable date by talking to initiator} will realize the goal \textbf{(Find) agreeable meeting date},
\item[] Which will promote the value \textbf{User friendliness}
\item[] \textit{Therefore} 
\item[] We should perform action \textbf{Find agreeable date by talking to initiator}
\end{itemize}

@@

These critical questions can point to new arguments that might counter the original argument. Take, for example, critical question 8. If new evidence comes up indicating that meeting schedulers do not care about user friendliness we have a counterargument to the above argument, `evidence shows that \textbf{user friendliness} is not a legitimate value'. Another way to counter an argument for an action is to suggest an alternative action that realizes the same goal (question 2) or an alternative goal that promotes the same value (question 3). For example, we can argue that \textbf{Find agreeable date using scheduler} also realizes the goal \textbf{Find agreeable meeting date}, which gives us a counterargument to the original argument -- to find a date by talking to the initiator -- that also follows PRAS.

In argumentation, counterarguments are said to \emph{attack} the original arguments (and sometimes vice versa). In the work of Atkinson et al.~\cite{atkinson2007}, arguments and their attacks are captured as an \emph{argumentation framework} of arguments and attack relations as introduced by Dung~\cite{Dung1995}\footnote{Full definitions of Dung's~\cite{Dung1995} frameworks and semantics will be given in section \ref{sect:gmas}. In this section, we will briefly discuss the intuitions behind these semantics.}. Figure \ref{fig:pras:example1} shows an argumentation framework with three arguments from the above example: the argument for \textbf{Find agreeable date by talking to initiator} (A1), the argument for \textbf{Find agreeable date using scheduler} (A3), and the argument that \textbf{User friendliness} is not a legitimate value (A2). The two alternative PRAS instantiations are A1 and A3. These arguments mutually attack each other, as \textbf{Find agreeable date by talking to initiator}  and \textbf{Find agreeable date using scheduler} are considered to be mutually exclusive. Argument A2 attacks A1, as it questions the legitimacy of the value \textbf{User friendliness}.

@@

In some cases, it is more difficult to determine whether or not an argument is defeated. Take, for example, the argumentation framework with just A1 and A3: they attack each other, they are alternatives and without any explicit preference, it is impossible to choose between the two. It is, however, possible to include explicit preferences between arguments when determining argument acceptability \cite{amgoud2002reasoning}. Take, for example, A1 and A3. If we say that we prefer the action \textbf{Find agreeable date by talking to initiator} (A1) over the action \textbf{Find agreeable date using scheduler} (A3), we remove the attack from A1 to A3 (Figure~\ref{fig:pras:example2}, left). This makes A1 the only undefeated argument, whereas A3 is now defeated. It is also possible to give explicit arguments for preferences \cite{modgil2009}. These arguments are then essentially attacks on attacks. For example, say we prefer A1 over A3 because `the \textbf{Meeting participant} does not like working with a scheduler' (A4). This can be rendered as a separate argument that attacks the attack from A3 to A1 (Figure~\ref{fig:pras:example2}, right), removing this attack and making $\{$A1, A4$\}$ the undefeated justified set of arguments.

>>>>

\subsection{Goal-oriented Requirements Language (GRL)}
\label{sect:background:grl}
GRL is a visual modeling language for specifying intentions, business goals, and non-functional requirements (NFRs) of multiple stakeholders \cite{Amyot:2010:EGM:1841349.1841356}. It can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions. A GRL model is a connected graph of intentional elements that optionally are part of actors. Figure~\ref{fig:trafficsim} illustrates a GRL diagram. An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system (\textbf{User}, Figure~\ref{fig:trafficsim}), or the system itself (\textbf{Simulation}, Figure~\ref{fig:trafficsim}). Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied. Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way. Softgoals (e.g. \textbf{Realistic simulation)} are often more related to NFRs, whereas goals (such as \textbf{Generate traffic}) are more related to functional requirements. Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals. In Figure~\ref{fig:trafficsim}, some of the tasks are \textbf{Create new cars} and \textbf{Keep same cars}. In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g. \textbf{******})

@@

Different links connect the elements in a GRL model. AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements. In Figure~\ref{fig:trafficsim}, the goal \textbf{Generate traffic} is XOR-decomposed to the tasks \textbf{Create new cars} and \textbf{Keep same cars}. Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element. A contribution link has a qualitative contribution type or a quantitative contribution. Task \textbf{Create new cars} has a \textbf{help} qualitative contribution to the softgoal \textbf{Realistic simulation}. Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors. For example, actor \textbf{Simulation} depends on the actor \textbf{User} to perform the task \textbf{Adjust car spawning rate} to fulfil its task \textbf{Create new cars}.

@@

In this paper, we use for our examples materials from our case study on the architecture for a traffic simulator \cite{}. \todo{Floris}{all}{briefly explain the traffic simulation case and the GRL model}

@@

\todo{Floris}{all}{These have to be amended}
\begin{itemize}
	\item Why are only the two softgoals \textbf{Quick} and \textbf{Low effort} selected for the task \textbf{Organize Meeting}? Why is, for example, a goal \textbf{Selecting the most convenient time} not included in the analysis for the actor \textbf{Meeting Initiator}?
	\item What does \textbf{Quality of Proposed Date} mean?
	\item How can one decide if the quality is low or high? Who is in charge? \todo{Floris}{Sepideh}{do you mean the \textbf{Quality of Proposed Date}? Or quality of the GRL diagram?}
	\item Does the meeting initiator use an email to inform the meeting participants to enter their availabilities? 
	\item Does the meeting scheduler uses an online calendar to obtain the available dates or an online website?
	\item How much time do the participants have to enter their availabilities? Is that important? Do they get a reminder from either the initiator or the scheduler? Is the reminder sent via an email? 
	\item What does \textbf{Richer Medium} mean? 
\end{itemize}

These are the type of the questions that we cannot answer just by looking at the GRL models. The model in figure~\ref{fig:trafficsim} does not contain information about discussions that let up to the resulting elements of the model, such as by various clarification steps for the naming, or alternatives that have been considered for the relationships. In order to address this, we use Argument Schemes for Goal Modeling (GMAS). GMAS can help us decide what intentional elements should remain in GRL and what new ones to be added or deleted, provide rationale behind the design decisions and the relationships between the links.

@@

So, for example, we can say that 
\begin{itemize}
\item[] \textbf{Generate traffic} is a goal,
\item[] \textbf{Keep same cars} will realize the goal \textbf{Generate traffic},
\item[] Which will promote the value \textbf{Simple design}
\item[] \textit{Therefore} 
\item[] We should perform action \textbf{Keep same cars}
\end{itemize}

@@

These critical questions can point to new arguments that might counter the original argument. Take, for example, critical question 4: if we find that \textbf{Keep same cars} actually negatively influences the value \textbf{Realistic simulation}, we have a counterargument to the above argument. Another way to counter an argument for an action is to suggest an alternative action that realizes the same goal (question 2) or an alternative goal that promotes the same value (question 3). For example, we can argue that \textbf{Create new cars} also realizes the goal \textbf{Generate traffic}, which gives us a counterargument to the original argument -- to generate traffic by simply keeping the cars that disappear off the screen and have them wrap around to the other side of the screen -- that also follows PRAS.

In argumentation, counterarguments are said to \emph{attack} the original arguments (and sometimes vice versa). In the work of Atkinson et al.~\cite{atkinson2007}, arguments and their attacks are captured as an \emph{argumentation framework} of arguments and attack relations as introduced by Dung~\cite{Dung1995}\footnote{Full definitions of Dung's~\cite{Dung1995} frameworks and semantics will be given in section \ref{sect:gmas}. In this section, we will briefly discuss the intuitions behind these semantics.}. Figure \ref{fig:pras:example1} shows an argumentation framework with three arguments from the above example: the argument for \textbf{Keep same cars} (A1), the argument for \textbf{Create new cars} (A3), and the argument that \textbf{Keep same cars} demotes the value \textbf{Realistic simulation} (A2). The two alternative PRAS instantiations are A1 and A3. These arguments mutually attack each other, as \textbf{Keep same cars}  and \textbf{Create new cars} are considered to be mutually exclusive. Argument A2 attacks A1, as it points to a negative side-effect of \textbf{Keep same cars}.

@@

In some cases, it is more difficult to determine whether or not an argument is defeated. Take, for example, the argumentation framework with just A1 and A3: they attack each other, they are alternatives and without any explicit preference, it is impossible to choose between the two. It is, however, possible to include explicit preferences between arguments when determining argument acceptability \cite{amgoud2002reasoning}. Take, for example, A1 and A3. If we say that we prefer the action \textbf{Create new cars} (A3) over the action \textbf{Keep same cars} (A1), we remove the attack from A1 to A3 (Figure~\ref{fig:pras:example2}, left). This makes A3 the only undefeated argument, whereas A1 is now defeated. It is also possible to give explicit arguments for preferences \cite{modgil2009}. These arguments are then essentially attacks on attacks. For example, say we prefer A3 over A1 because `it is important to have realistic traffic flows' (A4). This can be rendered as a separate argument that attacks the attack from A1 to A3 (Figure~\ref{fig:pras:example2}, right), removing this attack and making $\{$A3, A4$\}$ the undefeated set of arguments.
